An evaluation framework for trajectory inference methods

### 2017-11-24 analysis/paper/paper.Rmd
add back the exp_id parameter!!!!!

```{r, fig.width = 15, fig.height = 10}
n_methods_over_time <- readRDS(figure_file("n_methods_over_time.rds", experiment_name="method_characteristics"))
trajectory_components_over_time <- readRDS(figure_file("trajectory_components_over_time.rds", experiment_name="method_characteristics"))

@@

Given[d] this plethora of available trajectory inference methods, it is important that methods are compared, so that users can use the most optimal method available for their problem. Moreover, new methods need to be rigorously tested, so that development can focus on improving the current state-of-the-art. In this study, we therefore for the first time developed a comprehensive evaluation framework for trajectory inference methods. We test both on synthetic data, for which the gold standard is known, and real data, for which in some cases the gold standard can be extracted from current knowledge. We include a rigorous parameter optimisation. To make our framework reusable and extendable, we make use of continuous analysis [@beaulieu-jones_reproducibility_2017], which allows the developers of new methods to easily test their methods and compare them against the state-of-the-art.

# Results
## Evaluation workflow
### Method characterization
```{r}
# Load method characteristics
method_df <- readRDS(derived_file("method_df.rds", experiment_name="method_characteristics"))
```

@@

```{r, fig.width = 12, fig.height = 9}
method_characteristics <- readRDS(figure_file("method_characteristics.rds", experiment_name="method_characteristics"))

@@

## Overall method comparison
*Points to maybe raise*:
- Some methods can handle different trajectories which are not directly described in the main paper: eg. circular trajectories (for a lot of principal curves algorithms using `periodic.lowess`), trees (for a lot of bifurcation methods)

>>>>

```{r, fig.width = 15, fig.height = 10}
n_methods_over_time <- readRDS(figure_file("n_methods_over_time.rds", exp_id="method_characteristics"))
trajectory_components_over_time <- readRDS(figure_file("trajectory_components_over_time.rds", exp_id="method_characteristics"))

@@

Given[d] this plethora of available trajectory inference methods, it is important that methods are compared, so that users can use the most optimal method available for their problem. Moreover, new methods need to be rigorously tested, so that development can focus on improving the current state-of-the-art. In this study, we therefore for the first time developed a comprehensive evaluation framework for trajectory inference methods. We test both on synthetic data, for which the gold standard is known, and real data, for which in some cases the gold standard can be extracted from expert knowledge. We include a rigorous parameter optimisation, making sure that parameters are optimized while avoiding overfitting on characteristics of specific datasets. To make our framework reusable and extendable, we make use of continuous analysis [@beaulieu-jones_reproducibility_2017], which allows the developers of new methods to easily test their methods and compare them against the state-of-the-art.
# Results
## Evaluation workflow
### Method characterization
```{r}
# Load method characteristics
method_df <- readRDS(derived_file("method_df.rds", exp_id="method_characteristics"))
```

@@

```{r, fig.width = 12, fig.height = 9}
method_characteristics <- readRDS(figure_file("method_characteristics.rds", exp_id="method_characteristics"))

@@

## Overall method comparison
- Some methods can handle different trajectories which are not directly described in the main paper: eg. circular trajectories (for a lot of principal curves algorithms using `periodic.lowess`), trees (for a lot of bifurcation methods)
